### 2016-10-28 capability-machine.tex
comments

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
\subsubsection{Simple use of local capabilities}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv(l);
          l := 1;
          adv(0);
          return *l
\end{verbatim}
In this example \texttt{let l = 1 in} allocates a new local capability \texttt{l} with read-write permissions. Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
\lars{So this means that the value returned by *l is 1, correct ? We
  could add an assert statement to the example to show it more directly.}

\lars{What about an even simpler example, such as the one below:}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}

\subsection{Simple well-bracketness/local capability example}
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          adv();
          assert(x == 1);
          x := 2;
          adv();
          return 0
\end{verbatim}
A sensible calling scheme uses local capabilities for the continuation capability passed to the callee. In the above example, if \texttt{adv} can store the continuation, then 
\texttt{adv} can invoke it during the second call and thus force the assertion to fail.
\lars{don't you want to allocate the location \texttt{x} first and then
  return a function taking \texttt{adv} as an argument ??}

\subsection{Another simple well-bracketness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
\begin{verbatim}
let f = fun adv =>
          let l = 0 in
          let x = 1 in
          adv(l);
          assert(x == 1);
          x := 2;
          adv(0);
          return 0
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

@@

Standard region for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}

@@

\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

@@

Now define the value relation as follows:
\begin{align*}
  \stdvr &\defeq \lambda \; V \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(W))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\var{regions})(\start,\addrend,W) \\
             & \quad \text{for }\var{regions} = \dom(\erase{W}{\temp}) \union \dom(\erase{W}{\revoked})\}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \}
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwlx,\local)]} \mid \}
           \end{aligned}
\end{align*}

>>>>

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
\subsubsection{Simple use of local capabilities}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv(l);
          l := 1;
          adv(0);
          assert(l == 1)
\end{verbatim}
In this example \texttt{let l = 1 in} allocates a new local capability \texttt{l} with read-write permissions. Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
\lars{So this means that the value returned by *l is 1, correct ? We
  could add an assert statement to the example to show it more directly.}
\lau{Yes, and done.}

\lars{What about an even simpler example, such as the one below:}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
\lau{This example should certainly also hold true work, but it does not need local capabilities (it might serve well as a sanity check). }

\subsection{Simple well-bracketedness/local capability example}
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          adv();
          assert(x == 1);
          x := 2;
          adv();
          return 0
\end{verbatim}
A sensible calling scheme uses local capabilities for the continuation capability passed to the callee. In the above example, if \texttt{adv} can store the continuation, then 
\texttt{adv} can invoke it during the second call and thus force the assertion to fail.
\lars{don't you want to allocate the location \texttt{x} first and then
  return a function taking \texttt{adv} as an argument ??}
\lau{I am not sure I get what you have in mind, could you perhaps write it?}

\subsection{Another simple well-bracketedness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
\begin{verbatim}
let f = fun adv =>
          let l = 0 in
          let x = 1 in
          adv(l);
          assert(x == 1);
          x := 2;
          adv(0);
          return 0
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

@@

Standard region for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}

@@

\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

@@

Now define the value relation as follows:
\begin{align*}
  \stdvr &\defeq \lambda \; W \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(W))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\var{regions})(\start,\addrend,W) \\
             & \quad \text{for }\var{regions} = \dom(\erase{W}{\temp}) \union \dom(\erase{W}{\revoked})\}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \}
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwlx,\local)]} \mid \}
           \end{aligned}
\end{align*}
